
Analyzing the Tps6 samples, testing parameters in negative ion mode. Note that this comes
after playing around with the parameters with the vent DOM samples for a while
Krista Longnecker
April 9, 2014
========================================================

Keeping track of details by running this code as an R Markdown document. Essentially the code will be combined with the output into an HTML file. Note that there are three places where you will have to change
things based on which ionization mode you are running

1. I found a lower ppm in the centWave step was appropriate for negative ion mode data
2. I also pre-thresholded the data in negative ion mode to ignore peaks < 1000 (absolute intensity)
3. Around line 139 you need to decide which ionization mode you want at the end
4. In CAMERA, you need to select the ionization mode before finding adducts (line 217)

Go ahead and load up some of the parameters:
```{r Settings, eval=TRUE}
library(xcms)
library(CAMERA)
library(vegan)
library(Heatplus)
library(gplots)
library(RColorBrewer)
library(ggplot2)
library(ggdendro)


```

```{r setup, echo = TRUE, eval=TRUE}

knitr::opts_chunk$set(error=TRUE)

```

Now go ahead and do the peak picking, or load in the existing file...peak picking can be time consuming and it easier to only do that once

```{r GettingData, eval = TRUE}

##if you want to save time and process shorter time section, can set the 
##scanrange here (min and mx time, in seconds), and add scanrange = scanrange
##to the centWave peak picking below
#scanrange=c(100,300)
##Note: in negative ion mode, I have been setting a threshold for peaks in MSconvert before
##I even get to this step. This massively improves the peak picking in XCMS
usePath <- "C:/Users/Claire/Documents/MS_proc/UW_metabolomics_2015-08-18/files_to_process/2015-09-09.neg_batch_5percent"
ext <- ".mzML"
pre <- paste0(usePath,"/")
mzdatafiles <- list.files(usePath, recursive=FALSE, full.names=TRUE, pattern=paste0("\\", ext))
# mzdatafiles is class "character"

if (FALSE) {
## How many CPU cores has your machine (or cluster) ?
nSlaves=4

pdf(file = "test.pdf") #if want PDF file, add this after mzdiff: ,sleep = 0.0001 and uncomment dev.off()

#For negative ion mode: ppm = 2 seems best
#For positive ion mode: ppm = 3 seems best
xs<-xcmsSet(mzdatafiles, method = "centWave",ppm = 2 ,snthresh = 10,
  prefilter = c(5,1000), mzCenterFun = "wMean",integrate = 2, 
  verbose.columns = TRUE, peakwidth=c(20,60), fitgauss= TRUE, noise = 1000, 
  mzdiff=-0.005,nSlaves=nSlaves, sleep = 0.00001)

#scanrange = scanrange)
dev.off()

} else { 
  #load in a previous version of peak picking (picked with the above parameters)
  load("C:/Users/Claire/Dropbox (MIT)/MIT/alm_lab/UW_metabolomics_proc_shared/alignment_results/underworlds_24hr.2015-08-18-run.neg_batch_5percent.neg.proc-2015-09-09.Rimage")

  }



#setup the filenames all in one place bc easier...
fName_allPeaks <- "test.allpeaks.csv"
fName_aligned <- "test.aligned.csv"
NameOfFile <- "test.Rimage"

save.image(NameOfFile)

# #####the peakShape code seems to be from Tony Larson...via the Google Groups forum  
# ####turn this off 10/21/2013 (after v1 attempt)
#there is some sort of issue running this within r Markdown, fix it later
# source("removingNoise.R")  
# xs2 <- peakShape(xs,cor.val=0.9)


```

Setup the ability to rename the classes based on information in the CSV file

```{r getSampleInformation, eval=FALSE}
#figure out how to pass values to/from functions later...use this to match the files 
#with the information about what sample was run for each file
wDir <- "C:/Users/Claire/Dropbox (MIT)/MIT/alm_lab/UW_metabolomics_proc_shared/alignment_results/"
all <- read.csv(paste0(wDir,"underworlds_24hr.2015-08-18-run.neg_batch_5percent.neg.proc-2015-09-09.sequence_file.csv"))
putDataHere <-data.frame()

for(i in 1:length(all$File.Name)) {
  h <- all$File.Name[i]
  #paste0 will strip out the spaces; make the full filename here    
  h <-paste0(pre,h,ext)
  
  #this will pick up the right row in mzdatafiles
  m <- match(h,mzdatafiles)
  if (!is.na(m)) {
    tData <- all[i,]
    if (nrow(putDataHere)==0) {
      putDataHere <- tData
      } else {
    putDataHere[m,] <-tData
    }
    rm(tData)
  }   
  rm(h,m)
}   

rm(all) #keep this clean...

# rm(classlist) 
classlist <- matrix()
#this next bit is a hack, but it works in the context of xcms
for (i in 1:nrow(putDataHere)) {
  classlist[i] <-paste0(putDataHere$ionMode[i],".",putDataHere$GoodData[i])
}
rm(i)


```


```{r DoGrouping_1, eval=FALSE}
     
##now I will need to consider the groups. First 'group' is actually the ionization mode
#for simplicity, start by using split to get the files from one ionization mode  

# xs is what comes out of the xmcs() peak picking code
useXS <- xs

#define the classes after the peak picking...
#classlist was defined above, it's essentially a list of negative.1, negative.0, positive.1, postive.0 strings, corresponding to each sample

# sampclass(useXS) <- classlist adds (or replaces) the 'class' column in useXS with what's in classlist
sampclass(useXS) <- classlist 

#note this will select files in negative ion mode that are marked with '1' in the GoodData column

# here is where the classlist is used to select the subset of data!
# split returns a list of subsets of useXS that you can access using simple brackets
# so we want split(useXS, classlist)[class_of_interest]
xsOne <- split(useXS,classlist)[["negative.1"]]
putDataHere.1 <- split(putDataHere,classlist)[["negative.1"]]
nSamples<-length(xsOne@filepaths)

classlist.1 <- matrix()
#this next bit is a hack, but it works in the context of xcms
# this part is not necessary.
for (i in 1:nrow(putDataHere.1)) {
  classlist.1[i] <-paste0(putDataHere.1$ionMode[i],".",putDataHere.1$GoodData[i])#,".",
#                        putDataHere.1$sampleType[i])
}

#set the groups to the QC/blank/samples before grouping:
# classlist.1 is the same thing as classlist. She had originally added another .sampleType suffix to the values in the list but commented it out. So it's the same thing
sampclass(xsOne) <-classlist.1

# basically the important part of this is to get xsOne to contain just the batch of samples we wanna align
rc.obi <-retcor.obiwarp(xsOne,plottype="deviation",profStep = 0.1,distFunc = "cor", 
                         gapInit = 0.3,gapExtend=0.4)    

xgN3 <-group.density(xsOne,minfrac=0,minsamp=1,bw=30,mzwid=0.001)     

# oh this is the part where it finds RT/mz regions where samples had no peaks, and integrates over that region to make up a peak regardless
# i think this is working off of the aligned peaks - yes! bc it makes no sense to fill in peaks from totally different samples that have totally differnet peaks
xgF <- fillPeaks(xgN3, method = "chrom")                              

#do some housecleaning bc this file is getting huge
rm(rc.obi,xgN3,useXS)

```

```{r doQuickClusterAnalysis, eval = FALSE}
dataOut <-peakTable(xgF)
v <- colnames(dataOut)
h <-grep("_tps6_",v)

#so, h will have the index into the columns with actual data...keep those for the NMS
fNMS <-data.frame(dataOut[,h])

td <-decostand(fNMS,method= "pa") #convert to presence absence
bc <-vegdist(t(td),binary=TRUE,method = "bray")
hc <-hclust(bc,method = "average") #can also try ward

p <- ggdendrogram(hc,rotate = TRUE)
plot(p)

```

Go ahead and spit out some variables to the folder

```{r SaveData, eval = FALSE}

# NOT SURE WHAT THIS DOES - what is xgF at this point???
# do I need to do this for every batch? One time for all samples of the same mode?

#this will spit out the list of all features identified
write.csv(xgF@peaks,file = fName_allPeaks)

#if you want a table with everything lined up, use this:
#generally better to wait until after CAMERA
#write.csv(peakTable(xgF),file = "Tps6_aligned.2014.03.28.v1.csv")

```

These are the pieces for CAMERA
      
```{r CAMERA, eval = FALSE}
#
#nSamples <- length(mzdatafiles)
#xsa<-xsAnnotate(xgF,sample = c(1:nSamples))
#xsa<-xsAnnotate(xgF,sample = 34) randomly try one sample for now
xsa<-xsAnnotate(xgF)

#group the features initially just by retention time
xsaF <-groupFWHM(xsa)

#figure out which features also have a matching 13C feature. Have to enter both 
#the relative error (ppm) and the absolute error (mzabs)
xsaFI <-findIsotopes(xsaF,ppm=1.5,mzabs = 0.0001,minfrac = 1/nSamples)

#now group by the correlations based on (1) intensity, (2) EIC, (3) isotopes...
xsaC <-groupCorr(xsaFI,cor_eic_th=0.75,pval=0.05, graphMethod="hcs",
 calcIso = TRUE, calcCiS = TRUE, calcCaS = FALSE)

#setup the file to also look for adducts, only go with the primary adducts for the moment
file <-system.file("rules/primary_adducts_neg.csv",package = "CAMERA")
rules <-read.csv(file)
an <-findAdducts(xsaC,polarity = "negative",rules=rules,ppm=1.5)

#do some housecleaning
rm(xsa,xsaF,xsaFI,xsaC)
```


And save the final steps

```{r lastSave, eval =FALSE}
#can plot some variables       
#windows() #brings up a box for plotting that is outside RStudio (works better here)
#plotEICs(an,pspec=2,maxlabel=5)
#plotPsSpectrum(xsaC,2,maxlabel=5)

#export the data to the CSV file
write.csv(getPeaklist(an),file=fName_aligned) 
save.image(NameOfFile)
#


```

